---
title: "Report"
author: "Pedro Belém, Rui Fonseca, Tiago Botelho"
date: "December 23, 2016"
output:
  pdf_document:
    fig_height: 6
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gdata)
library(ggplot2)
library(dplyr)
library(reshape2)
library(tidyr)
library(lubridate)
library(DMwR)
library(rpart.plot)
library(MASS)
```

## Data Pre processing
We begin by reading the data from the crime.xls file into a data frame
```{r}
data_path <- "./crime.xls"
info <- read.xls(data_path, sheet=1)
```
Then we removed all the instances with outlier dates. To calculate the outliers we considered a date as the number of days since the beginning of unix time.
```{r}
dates <- info$Date
n_days <- day(days(ymd(dates)))
info <- info[n_days >= quantile(n_days, .25) - 1.5*IQR(n_days) & n_days <= quantile(n_days, .75) + 1.5*IQR(n_days), ]
```

Then, we convert all the missing values to NA.
```{r}
info$BlockRange[info$BlockRange=='UNK'] <- NA
info$Type[info$Type == '-'] <- NA
info$Suffix[info$Suffix == '-'] <- NA
info$Offense.Type[info$Offense.Type == '1'] <- NA
```


## Data Visualisation

You can also embed plots, for example:

```{r pressure, echo=FALSE}
library(gdata)
library(ggplot2)
library(dplyr)
library(reshape2)
library(tidyr)
data_path <- "./crime.xls"
info <- read.xls(data_path, sheet=1)
info$BlockRange[info$BlockRange=='UNK'] <- NA
info$Type[info$Type == '-'] <- NA
info$Suffix[info$Suffix == '-'] <- NA

split <- strsplit(as.character(info$BlockRange), "-")
info$BlockRange <- order(sapply(split, "[", 1))

#number of crimes per street name
info.df <- tbl_df(info)
by_street <- group_by(info.df, StreetName)
count <- arrange(tally(by_street), desc(n))
#reorder porque o x ficava ordenado por ordem alfabetica e quero pelo n, e -n para ficar por ordem descendente.
ggplot(head(count), aes(x=reorder(StreetName,-n), y=n)) + geom_bar(stat="identity") + ggtitle("Distribution of crimes per StreetName")
```



```{r echo = FALSE}
#number of crimes per beat with the types of crimes
info.df <- tbl_df(info) %>% drop_na(Beat, BlockRange)
by_beat_with_group <- group_by(info.df, Beat, BlockRange)
by_beat <- group_by(info.df, Beat)
crime_count_with_group <- arrange(tally(by_beat_with_group), desc=-n)
crime_count <- arrange(tally(by_beat), desc=-n)
# we only want the top ones
top_beats <- head(crime_count)
top_crime_count_with_group <- crime_count_with_group[crime_count_with_group$Beat %in% top_beats$Beat,]
top_crime_count_with_group <- subset(crime_count_with_group, Beat %in% top_beats$Beat)
ggplot(top_crime_count_with_group, aes(reorder(Beat, -n), n, fill=BlockRange, order=BlockRange)) + geom_bar(stat="identity") + ggtitle("Distribution of crimes per beat")
```

##Data Prediction
The question that we're going to answer is: since we know all these crimes that happen, how many offenses will occur, in a given day interval of a day, in a certain police beat. First, we slit the hours of a day into three diferent intervals. The first one is the mourning, starts at 8h(inclusive) until 12h(exclusive). The second one is in the afternoon, starts at 12h (inclusive) until 19h (exclusive). The third interval (night) starts at 19h(exclusive) and ends at 8h (exclusive) of the next day (e.g. if a crime happened at 4th of day 20, it will correspond to the night period of the day 19).
We found that some columns of the original data aren't necessary for our predictive model. 
Only some information for this problem is actually needed so We had to adapt the given data for this problem. First we noticed that the premise of the crimes are irrelevant and that the only necessary attribute to identify an area is the beat. So our final data model for this problem consists of

DayInterval | Beat | Day | Month | Year | Offenses
--------------------------------------------------

To convert the given data to this data model, we start by adding the column with the correspondent day interval, and then we create a new data frame only with those column. As there are multiple crimes in the same day interval but we want the number of offenses of the day interval we sum the offenses on the same interval. 

```{r}

dataset_prep <- function(x) {
  x$Date <- ifelse(as.integer(x$Hour) < 8,
                   as.character(as.Date(x$Date) - 1), 
                   as.character(x$Date))
  
  # Split info in time intervals
  x$DayInterval <- 0
  x[as.integer(x$Hour) < 8 | as.integer(x$Hour) >= 19,]$DayInterval <- 3
  x[as.integer(x$Hour) >= 12 & as.integer(x$Hour) < 19,]$DayInterval <- 2
  x[as.integer(x$Hour) >= 8 & as.integer(x$Hour) < 12,]$DayInterval <- 1

  ret <- data.frame(WeekDay = as.integer(strftime(x$Date, "%u")),
                    DayInterval = x$DayInterval,
                    Beat = x$Beat,
                    Offenses = x$X..offenses,
                    Day = day(x$Date),
                    Month = month(x$Date),
                    Year = year(x$Date),
                    stringsAsFactors = FALSE)
  ret <- group_by(ret, WeekDay, DayInterval, Beat, Day, Month, Year) %>% 
    summarize(Offenses = sum(Offenses))
  
  return(ret)
}

```

At this moment we have the given data adjusted to the problem but we only have cases where was crimes, so we needed to generate all combinations with 0 offenses. 

So we generate a new data frame with all combinations of DayInterval, Beat, Day, Month, Year and added to our data only the combinations missing, with Offenses as 0. To do that we merged the 2 data frames and removed the rows with duplicated combination (DayInterval, Beat, Day, Month, Year) and 0 Offenses. 
```{r eval=FALSE}
get_days_between <- function(info) {
  firstdate <- info$Date[order(format(as.Date(info$Date)))[1]]
  lastdate <- info$Date[tail(order(format(as.Date(info$Date))), n=1)] 
  
  return(seq(as.Date(firstdate), as.Date(lastdate), by="days"))
}

create_total_perm <- function(preprocessed) {
  days.between <- get_days_between(info)
  unique_beats <- unique(preprocessed$Beat)
  unique_day_intervals <- unique(preprocessed$DayInterval)
  
  all_beats_perm <- data.frame(Date = rep(days.between, 
                                          times=length(unique_beats) * 
                                            length(unique_day_intervals)),
                               DayInterval = rep(unique_day_intervals, 
                                                 times=length(unique_beats) * 
                                                   length(days.between)),
                               Beat = rep(unique_beats, 
                                          times=length(unique_day_intervals) * 
                                            length(days.between)),
                               Offenses = rep(0, 
                                              times=length(unique_day_intervals) * 
                                                length(days.between) * 
                                                length(unique_beats)))
  all_beats_perm$DayInterval <- as.integer(all_beats_perm$DayInterval)
  all_beats_perm$Offenses <- as.character(all_beats_perm$Offenses)
  all_beats_perm$WeekDay = as.integer(strftime(all_beats_perm$Date, "%u"))  
  all_beats_perm$Day <- day(as.Date(all_beats_perm$Date))
  all_beats_perm$Month <- month(as.Date(all_beats_perm$Date))
  all_beats_perm$Year <- year(as.Date(all_beats_perm$Date))
  all_beats_perm <- all_beats_perm[,!colnames(all_beats_perm) %in% c("Date")]
  
  return(all_beats_perm)
}

info.preprocessed.total_perm <- create_total_perm(info.preprocessed)
info.preprocessed.joined <- merge(info.preprocessed, info.preprocessed.total_perm, 
                                  by=c("WeekDay", "DayInterval", "Beat", "Day", "Month", "Year"),
                                  all=TRUE)
info.preprocessed.joined[is.na(info.preprocessed.joined$Offenses.x),]$Offenses.x <- 0
info.preprocessed.joined$Offenses <- info.preprocessed.joined$Offenses.x
info.preprocessed.joined <- info.preprocessed.joined[,!colnames(info.preprocessed.joined) %in%
                                                       c("Offenses.y", "Offenses.x")]
```

We considered that the given data had all occorrences between the first and the last date of the set.\

At this moment we have all the data structured and we can 