---
title: "Report"
author: "Pedro Belém, Rui Fonseca, Tiago Botelho"
date: "December 23, 2016"
output:
  html_document: default
  pdf_document:
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gdata)
library(ggplot2)
library(dplyr)
library(reshape2)
library(tidyr)
library(lubridate)
library(DMwR)
library(rpart.plot)
library(MASS)
```

## Data Pre processing
The data is read from the crime.xls file into a data frame
```{r}
data_path <- "./crime.xls"
info <- read.xls(data_path, sheet=1)
```
We remove outliers by unix time days
```{r}
dates <- info$Date
n_days <- day(days(ymd(dates)))
info <- info[n_days >= quantile(n_days, .25) - 1.5*IQR(n_days) & n_days <= quantile(n_days, .75) + 1.5*IQR(n_days), ]
```

Then, the unknown values (NA) are treated
```{r}
info$BlockRange[info$BlockRange=='UNK'] <- NA
info$Type[info$Type == '-'] <- NA
info$Suffix[info$Suffix == '-'] <- NA
info$Offense.Type[info$Offense.Type == '1'] <- NA
```
and the block range variable is modified. Since it is always a string like "X-Y", with X being a multiple of 100 and Y=X+99, we keep only X/100
```{r}
split <- strsplit(as.character(info$BlockRange), "-")
info$BlockRange <- order(sapply(split, "[", 1))
```

## Data Visualisation

You can also embed plots, for example:

```{r pressure, echo=FALSE}
library(gdata)
library(ggplot2)
library(dplyr)
library(reshape2)
library(tidyr)
data_path <- "./crime.xls"
info <- read.xls(data_path, sheet=1)
info$BlockRange[info$BlockRange=='UNK'] <- NA
info$Type[info$Type == '-'] <- NA
info$Suffix[info$Suffix == '-'] <- NA

split <- strsplit(as.character(info$BlockRange), "-")
info$BlockRange <- order(sapply(split, "[", 1))

#number of crimes per street name
info.df <- tbl_df(info)
by_street <- group_by(info.df, StreetName)
count <- arrange(tally(by_street), desc(n))
#reorder porque o x ficava ordenado por ordem alfabetica e quero pelo n, e -n para ficar por ordem descendente.
ggplot(head(count), aes(x=reorder(StreetName,-n), y=n)) + geom_bar(stat="identity") + ggtitle("Distribution of crimes per StreetName")
```



```{r echo = FALSE}
#number of crimes per beat with the types of crimes
info.df <- tbl_df(info) %>% drop_na(Beat, BlockRange)
by_beat_with_group <- group_by(info.df, Beat, BlockRange)
by_beat <- group_by(info.df, Beat)
crime_count_with_group <- arrange(tally(by_beat_with_group), desc=-n)
crime_count <- arrange(tally(by_beat), desc=-n)
# we only want the top ones
top_beats <- head(crime_count)
top_crime_count_with_group <- crime_count_with_group[crime_count_with_group$Beat %in% top_beats$Beat,]
top_crime_count_with_group <- subset(crime_count_with_group, Beat %in% top_beats$Beat)
ggplot(top_crime_count_with_group, aes(reorder(Beat, -n), n, fill=BlockRange, order=BlockRange)) + geom_bar(stat="identity") + ggtitle("Distribution of crimes per beat")
```

##Data Prediction
The question that we're going to answer is: since we know all these crimes that happen, how many offenses will occur, in a given day interval of a day, in a certain police beat. First, we slit the hours of a day into three diferent intervals. The first one is the mourning, starts at 8h(inclusive) until 12h(exclusive). The second one is in the afternoon, starts at 12h (inclusive) until 19h (exclusive). The third interval (night) starts at 19h(exclusive) and ends at 8h (exclusive) of the next day (e.g. if a crime happened at 4h of day 20, it will correspond to the night period of the day 19).
