---
title: "Report"
author: "Pedro Belém, Rui Fonseca, Tiago Botelho"
date: "December 23, 2016"
output:
  pdf_document:
    fig_height: 6
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(gdata)
library(ggplot2)
library(dplyr)
library(reshape2)
library(tidyr)
library(lubridate)
library(DMwR)
library(rpart.plot)
library(MASS)
```



## Data Pre processing
The data is read from the crime.xls file into a data frame
```{r}
data_path <- "./crime.xls"
info <- read.xls(data_path, sheet=1)
```

Next we remove outliers by unix time days
```{r}
remove_outliers <- function(info) {
  n_days <- day(days(ymd(info$Date)))
  return(info[n_days >= quantile(n_days, .25) - 1.5*IQR(n_days) & n_days <= quantile(n_days, .75) + 1.5*IQR(n_days),])
}
```

Then, the unknown values (NA) are treated
```{r}
na_handler <- function(info) {
  info$Beat[info$Beat == 'UNK'] <- NA
  info <- knnImputation(info, k=3)
  info$BlockRange[info$BlockRange=='UNK'] <- NA
  info$Type[info$Type == '-'] <- NA
  info$Suffix[info$Suffix == '-'] <- NA
  info$Beat <- as.character(info$Beat)
  
  return(info)
}
```

Tthe block range variable is modified. Since it is always a string like "X-Y", with X being a multiple of 100 and Y=X+99, we keep only X/100
```{r}
split <- strsplit(as.character(info$BlockRange), "-")
info$BlockRange <- order(sapply(split, "[", 1))
```


We will also create a data frame that contains for each crime:

* WeekDay: the day of the week when the crime happene
* Beat: the police beat
* DayInterval: the interval of the day in which the crime happened, following this criteria:
    1. "Mourning" represented as "1", is from the hour interval 8 >= h > 12
    2. "Afternoon" represented as "2", is from the hour interval 12 >= h > 19
    3. "Night"represented as "3", is from the hour interval 19 >= h <= 23 of the same day, plus the hour interval 0 >= h > 8 of the following day
* Year: the year in which the crime happened
* Month: the month in which the crime occoured
* Day: the day of the month in which the crime happened


Create the data frame with the new collumns
```{r}
dataset_prep <- function(x, only.week=FALSE) {
  x$Date <- ifelse(as.integer(x$Hour) < 8, as.character(as.Date(x$Date) - 1), as.character(x$Date))
  
  # Split info in time intervals
  x$DayInterval <- 0
  x[as.integer(x$Hour) < 8 | as.integer(x$Hour) >= 19,]$DayInterval <- 3
  x[as.integer(x$Hour) >= 12 & as.integer(x$Hour) < 19,]$DayInterval <- 2
  x[as.integer(x$Hour) >= 8 & as.integer(x$Hour) < 12,]$DayInterval <- 1

  if(only.week){
    return(data.frame(WeekDay = as.integer(strftime(x$Date, "%u")),
                      DayInterval = x$DayInterval,
                      Beat = x$Beat,
                      Offenses = x$X..offenses,
                      stringsAsFactors = FALSE))
  }
  
  return(data.frame(WeekDay = as.integer(strftime(x$Date, "%u")),
                    DayInterval = x$DayInterval,
                    Beat = x$Beat,
                    Offenses = x$X..offenses,
                    Day = day(x$Date),
                    Month = month(x$Date),
                    Year = year(x$Date),
                    stringsAsFactors = FALSE))
}
```

```{r echo = FALSE}
get_days_between <- function(info) {
  firstdate <- info$Date[order(format(as.Date(info$Date)))[1]]
  lastdate <- info$Date[tail(order(format(as.Date(info$Date))), n=1)] 
  
  return(seq(as.Date(firstdate), as.Date(lastdate), by="days"))
}
```

Create all permutations of missing 
```{r}
create_total_perm <- function(preprocessed, only.week = FALSE) {
  days.between <- get_days_between(info)
  unique_beats <- unique(preprocessed$Beat)
  unique_day_intervals <- unique(preprocessed$DayInterval)
  
  if(only.week){
    unique_weekdays <- unique(preprocessed$WeekDay)
    all_beats_perm <- data.frame(WeekDay = rep(unique_weekdays, times=length(unique_beats) * length(unique_day_intervals)),
                                 DayInterval = rep(unique_day_intervals, times=length(unique_beats) * length(days.between)),
                                 Beat = rep(unique_beats, times=length(unique_day_intervals) * length(days.between)),
                                 Offenses = rep(0, times=length(unique_day_intervals) * length(days.between) * length(unique_beats)))
  } else{
    all_beats_perm <- data.frame(Date = rep(days.between, times=length(unique_beats) * length(unique_day_intervals)),
                                 DayInterval = rep(unique_day_intervals, times=length(unique_beats) * length(days.between)),
                                 Beat = rep(unique_beats, times=length(unique_day_intervals) * length(days.between)),
                                 Offenses = rep(0, times=length(unique_day_intervals) * length(days.between) * length(unique_beats)))
    all_beats_perm$DayInterval <- as.integer(all_beats_perm$DayInterval)
    all_beats_perm$Beat <- as.numeric(all_beats_perm$Beat)
    all_beats_perm$Offenses <- as.character(all_beats_perm$Offenses)
    all_beats_perm$WeekDay = as.integer(strftime(all_beats_perm$Date, "%u"))  
    all_beats_perm$Day <- day(as.Date(all_beats_perm$Date))
    all_beats_perm$Month <- month(as.Date(all_beats_perm$Date))
    all_beats_perm$Year <- year(as.Date(all_beats_perm$Date))
    all_beats_perm <- all_beats_perm[,!colnames(all_beats_perm) %in% c("Date")]
  }
  
  return(all_beats_perm)
}

```


## Data Visualisation

```{r echo=FALSE}
#this is the variable initialization
info <- read.xls(data_path, sheet=1) %>% remove_outliers %>% na_handler
split <- strsplit(as.character(info$BlockRange), "-")
info$BlockRange <- sapply(split, "[", 1)

info.preprocessed <- dataset_prep(info)
info.preprocessed.group <- group_by(info.preprocessed, WeekDay, DayInterval, Beat, Day, Month, Year) %>% summarize(Offenses = sum(Offenses))
info.preprocessed.total_perm <- create_total_perm(info.preprocessed)
info.preprocessed.joined <- merge(x=info.preprocessed.total_perm, y=info.preprocessed.group, all=TRUE)


info.preprocessed.onlyweek <- dataset_prep(info, only.week = TRUE)
info.preprocessed.onlyweek.group <- group_by(info.preprocessed.onlyweek, WeekDay, DayInterval, Beat) %>% summarize(Offenses = sum(Offenses))
info.preprocessed.onlyweek.total_perm <- create_total_perm(info.preprocessed, only.week = TRUE)
```

Creating graphics with the data, we can begin to find some usefull patterns in the data.

```{r echo=FALSE}

#number of crimes per beat with the types of crimes
info.df <- tbl_df(info) %>% drop_na(Beat, BlockRange)
by_beat <- group_by(info.df, Beat)
crime_count <- arrange(tally(by_beat), desc(n))
# we only want the top ones
top_beats <- head(crime_count)
top_beat_crimes <- filter(info.df,info.df$Beat %in% top_beats$Beat)
ggplot(top_beat_crimes, aes(x = Beat, fill=BlockRange)) + geom_bar(stat="count") + ggtitle("Distribution of crimes per beat")

```

```{r echo=FALSE}
#number of crimes per Address (block range, streetname, type, suffix)
info.df <- tbl_df(info)
by_address <- group_by(info.df,BlockRange, StreetName, Type, Suffix)
count <- arrange(tally(by_address), desc(n))
#criar a morada com as outras colunas coladas
count$d <- paste(count$BlockRange, count$StreetName,count$Type,count$Suffix)
#remover os NA's depois, antes tava dificil
count$d <- gsub(" NA","",count$d)
count$d <- gsub("NA ","",count$d)
#reorder porque o x ficava ordenado por ordem alfabetica e quero pelo n, e -n para ficar por ordem descendente.
ggplot(head(count), aes(x=reorder(d,-n), y=n)) + geom_bar(stat="identity") + ggtitle("Distribution of crimes per Address")
```

```{r echo=FALSE}
#number of crimes per street name, Type and Suffix
info.df <- tbl_df(info)
by_street_type_suffix <- group_by(info.df, StreetName, Type, Suffix)
count <- arrange(tally(by_street_type_suffix), desc(n))
#criar a morada com as outras colunas coladas
count$d <- paste(count$StreetName,count$Type,count$Suffix)
#remover os NA's depois, antes tava dificil
count$d <- gsub(" NA","",count$d)
#reorder porque o x ficava ordenado por ordem alfabetica e quero pelo n, e -n para ficar por ordem descendente.
ggplot(head(count), aes(x=reorder(d,-n), y=n)) + geom_bar(stat="identity") + ggtitle("Distribution of crimes per Address except BlockRange")
```

```{r echo=FALSE}
#number of crimes per street name
info.df <- tbl_df(info)
by_street <- group_by(info.df, StreetName)
count <- arrange(tally(by_street), desc(n))
#reorder porque o x ficava ordenado por ordem alfabetica e quero pelo n, e -n para ficar por ordem descendente.
ggplot(head(count), aes(x=reorder(StreetName,-n), y=n)) + geom_bar(stat="identity") + ggtitle("Distribution of crimes per StreetName")
```


```{r echo = FALSE}
#number of crimes per hour and type
ggplot(info, aes(x=Hour, color=Offense.Type)) + geom_histogram(binwidth = 1) + ggtitle("Distribution of crimes per hour and type")

```

```{r echo=FALSE}
#number of crimes per hour and type
ggplot(info, aes(x=Hour)) + geom_histogram(binwidth = 1) + facet_wrap(~ Offense.Type) + ggtitle("Distribution of crimes per hour and type")

```

##Data Prediction
The question that we're going to answer is: since we know all these crimes that happen, how many offenses will occur, in a given day interval of a day, in a certain police beat. 
